Rewind: Automatically Generated\\Digital Memories from Mobile Geolocation Data

### 2016-09-20 paper/uist_2016/paper.tex

% heather - need to be rephrased perhaps? Should we give a high level summary of all the relationships we've found? More analysis to be done for photo vs. videos, another table?
% semantic vs episodic memory
\section{Findings}
% upper-bound estimates of event-related memories.  <-- april - found this without the beginning of its sentence. where does this fragment belong?
The findings of our user study center on the participants’ relationship to the Rewinds generated from their location data compared to the static photos. We focus on the participants’ judgments of their Rewinds’ memorability, believability, and desirability in comparison to their judgments of the photos paired with the Rewinds. We also discuss the participants' interactions with and reactions to their Rewinds. The results use data from the user studies conducted with 12 men and 6 women. Half of the participants identified themselves as frequent travelers. Additionally, half of the participants were regular GPS users, which we defined as people who use GPS navigation even in locations already familiar to them.

@@

\subsection{Within-Subject Study}
\subsubsection{Memorability}
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with Rewinds than with static street-view photos, $\chi^2(1, 356) = 96.987$, $p < 0.01$. When being presented with a street-side photo, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants played accompanying Rewinds. It should not be too surprising that cinematic media like a Rewind would provide participants with more contextual meanings and retrival cues for semantic processing, as more than one participant have expressed their contentment after wathing \textit{Rewind} videos by claiming ``I recognize the photo now!'' when their initial attempts with static images were fruitless.

It could possibly be inferred that participants are likely to benefit more from \textit{Rewind} videos than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip remembrance when presented with static images, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip. %- tbc
\subsubsection{Believability and Desirability}
Although not statistically significant, %TODO TODO TODO TODO TODO

\subsection{Between-Subject Study}
\begin{table*}[t]
	\centering
	\caption{My caption}
	\small
	\label{within-subject-table}
	\begin{tabular}{c|c|c|c|c|c|c|c|c|}
		\cline{2-9}
		& \multicolumn{4}{c|}{Memorability} & \multicolumn{2}{c|}{Belivability} & \multicolumn{2}{c|}{Desirability} \\ \cline{2-9} 
		& \begin{tabular}[c]{@{}c@{}}video\\  location\end{tabular} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}photo\\ location/trip\end{tabular}} & \begin{tabular}[c]{@{}c@{}}video\\ trip\end{tabular} & photo & video & photo & video \\ \hline
		\multicolumn{1}{|c|}{N} & 177 & \multicolumn{2}{c|}{179} & 177 & 86 & 128 & 179 & 177 \\ \hline
		\multicolumn{1}{|c|}{\%} & 96.61\% & \multicolumn{2}{c|}{49.72\%} & 53.67\% & 77.91\% & 79.69\% & 10.61\% & 14.69\% \\ \hline
		\multicolumn{1}{|c|}{p-val} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 96.987$\\ p = 0.0001\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 0.41$\\ p = 0.2611\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 214) = 0.02$\\ p = 0.4434\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356 ) = 0.995$\\ p = 0.1593\end{tabular}} \\ \hline
	\end{tabular}
\end{table*}
% 2nd table %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Believability}
Rewinds should be believable as memories, and be able to serve as digitally-sourced memories of locations and trips. For that purpose, we define the believability of a Rewind as its accuracy to the participant's memories. Here we will look at how GPS usage and visit frequency affect participants' believability judgments, and discuss the features that participants pointed out as either contributing to or detracting from a Rewind's believability.

@@

\subsection{Desirability}
For Rewinds to be useful, they need to have value to their owners. We measured the value of photos and Rewinds by asking participants if they would want to keep the photo or Rewind after seeing it. Overall, participants wished to keep 11\% of the photos they were shown and 15\% of the Rewinds they were shown. Because these were images from third-party services, they are naturally less likely to be desirable than media captured by the user themselves. In this section we will look at how the value of Rewind's digitally-fabricated memories differs between participants who consider themselves frequent travelers and participants who don't. We will also examine how visit frequency and travel companionship affect the desirability of a Rewind.

>>>>

% heather - need to be rephrased perhaps? Should we give a high level summary of all the relationships we've found? More analysis to be done for photo vs. videos, another table?
% semantic vs episodic memory
% mention distance categories in the beginning (up up up) + everything else 
\section{Findings}
The findings of our user study center on the participants’ relationship to the Rewinds generated from their location data compared to the static photos. We focus on the participants’ judgments of their Rewinds’ memorability, believability, and desirability in comparison to their judgments of the photos paired with the Rewinds. We also discuss the participants' interactions with and reactions to their Rewinds. The results use data from the user studies conducted with 12 men and 6 women. Half of the participants identified themselves as frequent travelers. Additionally, half of the participants were regular GPS users, which we defined as people who use GPS navigation even in locations already familiar to them.

@@

\subsection{Within-subject study}
\subsubsection{Semantic Memory} 
Within-subject comparison in our study reveals that participants are significantly better at recalling semantic memory of a particular location with \textit{Rewind} videos than with static street-view images, $\chi^2(1, N=356) = 96.987$, $p <.01$. When being presented with a street-side image retrieved based on their location data, participants were able to recognize places visited 50\% of the time; yet recognition rate of locations reached 97\% when participants started playing accompanied \textit{Rewind} video, with a significant increase of 47\%. It is perhaps not too surprising as cinematic media like \textit{Rewind} video would provide users with more contextual meanings and retrival cues for semantic processing, as three participants (P4, P6 and P10) have once expressed their contentment after watching \textit{Rewind} videos by claiming ``I recognize the photo now!'' when their initial attempts with static images recognition were fruitless.

%
\subsubsection{Episodic Memory}
It could possibly be inferred that participants are likely to benefit more from \textit{Rewind} videos than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip rememberance when presented with static images as opposed to what they were asked after watchiing \textit{Rewind} videos, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip; since if participants fail to form any cognitive association with the locations depicted in the images, it is very unlikely that they would be able to associate target itineraries with the locations shown. Thus, it is perhaps safe to presume that location rememberance with images bounds above the maximum likelihood of trip rememberance with the same images presented to the participants, and we used location rememberance from image recognition as an upper-limit gauge against trip rememberance from \textit{Rewind} videos thereafter.

% h - add another table, only abt memorability tho
% explain distance groups in the beginning
% where is home? need to explain this part in system?
Using the aforementioned presumption, our study examined effects of \textit{Rewind} videos on episodic processing in metric-specific groups, as well as the performance overall. Our study shows that comparing to static street-view images, recollection of event-driven memories using \textit{Rewind} videos is significantly improved for trips taken that were 100 - 1,000 km (62 - 620 miles) away from home by 50\%, $\chi^2(1, N=20) = 3.215$, $p <.05$; and for trips that were least frequently taken by about 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$. Our study also suggested that GPS users would benefit more from \textit{Rewind}, even though the difference was mildly significant by an increase about 15\%, $\chi^2(1, N=118) = 2.145$, $p = .0715 (<.1)$. 
% h - prob explain this a little bit more

One plausible explanation of the improvement may be inherent to the encoding of episodic memory in the first place. As suggested by the levels-of-processing theory(ref), encoding efficacy is heavily dependent on the levels to which stimuli are being actively or passively processed. Not surprisingly, superficial mental engagement would produce weaker, and less enduring cognitive representation which would demand more contextually accurate and abundant cues for memory triggery, and vice versa. Oftentimes, the cognitive representation resulted from shallow processing may not even be retrievable at all as it is simply not available for retrieval.  As previously described, both of the trips revealed in our results, those that were taken farthest away from home and those that were least frequently taken, may not be special enough to engage deep memory encoding nor often enough to reinforce incidental remembrance. This means that it would be more difficult for participants to recall these trips if the provided cues are not effective enough. Since \textit{Rewind} videos provide richer and more contextually meaningful cues than static street-side images, it is reasonable that \textit{Rewind} significantly outperforms plain images as suggested by our statistical analysis. While being asked about location and trip rememberance on one of his least frequently taken routes, P7 was not able to identify the location from the static image at first; yet he recognized the trip from \textit{Rewind} video as ``one street (in the video) helps me (him) remember". 
%The Effects of Divided Attention on Encoding and Retrieval Processes in Human Memory 
% http://www2.psych.ubc.ca/~pgraf/Psy583Readings/Craik%20et%20al%201996.pdf
% Craik et al. 1996

The extent to which attention is dedicated while taking a trip may be another possible explanation for the observed improvement using Rewinds. This is perhaps particularly applicable to explain why participants who were self-identified regular GPS users tend to benefit more from Rewinds with episodic memory recalling. Laboratory work done by Craik et al. 1996 (ref) has demonstrated that the encoding of memory is more effective under full-attention conditions than divided-attention conditions.  This means that regular GPS users, who are often directed step-by-step on how to navigate through their environments, may be less likely to engage their attention while taking their routes, leading to less available cognitive representation for retrieval. P3, a self-identified regular GPS user who had 30\% recalling rate using static images and 80\% trip remembrance rate using Rewinds said ``I really don't pay attention to my surroundings when I'm driving. (I am) Focused (focused) on GPS and road"  Under this circumstance, it is almost certain that Rewinds would provide a more successful access to episodic memory as it provides more spatial-temporal cues for memory retrieval, by not only adjusting image attributes to reflect timely and weather changes, but also mimicing the motions of the users when they travel down a particular path.

It may appear to be puzzling that participants seem to recall less about specifics of their trips from \textit{Rewind} videos compared to static images when they traveled alone, with a significant decrease around 40\%, $\chi^2(1, N=168) = 33.831$, $p <.01$; when the trips were most frequently taken, with a significant decrease of 20\%, $\chi^2(1, N=161) = 5.977$, $p <.01$; and when the trips taken were 10 - 100 meters (33 - 328 feet) away from home, with a mildly significant decrease of 50\%, $\chi^2(1, N=16) = 2.25$, $p <.1$.

It could possibly be inferred that participants are likely to benefit more from \textit{Rewind} videos than static street-view images at recalling episodic memory. Although participants were not explicitly asked about trip remembrance when presented with static images, location rememberance from static images may serve as an upper-bound estimation of how much an individual could recall about a particular trip. %- tbc

\subsubsection{Believability and Desirability}
Although not statistically significant, %TODO TODO TODO TODO TODO

\subsection{Between-Subject Study}
%%UPDATE BOTH TABLES AND ADD ONE MORE%%%
%%%%%%%%%%%percentage to decimal
\begin{table*}[t]
	\centering
	\caption{My caption}
	\small
	\label{within-subject-table}
	\begin{tabular}{c|c|c|c|c|c|c|c|c|}
		\cline{2-9}
		& \multicolumn{4}{c|}{Memorability} & \multicolumn{2}{c|}{Belivability} & \multicolumn{2}{c|}{Desirability} \\ \cline{2-9} 
		& \begin{tabular}[c]{@{}c@{}}video\\  location\end{tabular} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}photo\\ location/trip\end{tabular}} & \begin{tabular}[c]{@{}c@{}}video\\ trip\end{tabular} & photo & video & photo & video \\ \hline
		\multicolumn{1}{|c|}{N} & 177 & \multicolumn{2}{c|}{179} & 177 & 86 & 128 & 179 & 177 \\ \hline
		\multicolumn{1}{|c|}{\%} & 96.61\% & \multicolumn{2}{c|}{49.72\%} & 53.67\% & 77.91\% & 79.69\% & 10.61\% & 14.69\% \\ \hline
		\multicolumn{1}{|c|}{p-val} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 96.987$\\ p = 0.0001\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356) = 0.41$\\ p = 0.2611\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 214) = 0.02$\\ p = 0.4434\end{tabular}} & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}$\chi^2(1, 356 ) = 0.995$\\ p = 0.1593\end{tabular}} \\ \hline
	\end{tabular}
\end{table*}
% 2nd table %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Believability}
Rewinds should be believable as memories, and be able to serve as digitally-sourced memories of locations and trips. For that purpose, we define the believability of a Rewind as its accuracy to the participant's memories. Here we will look at how GPS usage and visit frequency affect participants' believability judgments, and discuss the features that participants pointed out as either contributing to or detracting from a Rewind's believability.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Desirability}
For Rewinds to be useful, they need to have value to their owners. We measured the value of photos and Rewinds by asking participants if they would want to keep the photo or Rewind after seeing it. Overall, participants wished to keep 11\% of the photos they were shown and 15\% of the Rewinds they were shown. Because these were images from third-party services, they are naturally less likely to be desirable than media captured by the user themselves. In this section we will look at how the value of Rewind's digitally-fabricated memories differs between participants who consider themselves frequent travelers and participants who don't. We will also examine how visit frequency and travel companionship affect the desirability of a Rewind.
